---
title: "IntersexDominance"
author: "Elise Huchard, Peter Kappeler, Claudia Fichtel, Nikos Smit, Dieter Lukas"
date: '`r Sys.Date()`'
output:
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
    code_folding: hide 
  github_document: 
    toc: true
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
  md_document: 
    toc: true
---


### This file contains the R script with the analyses reported in the manuscript:
The evolution of male-female dominance relations in primate societies
Authors: Elise Huchard, Peter M. Kappeler, Nikolaos Smit, Claudia Fichtel, Dieter Lukas
https://doi.org/10.1073/pnas.2500405122


# This script can be used to load the required data from the repository at http://github.com/dieterlukas/primate_power , run the analyses, and generate the figures.



```{r load data}



# Loading the relevant packages for tree manipulation (ape/geiger), data wrangling (dplyr), 
# and Bayesian analyses (rethinking/cmdstanr). To install these latter two packages, please see: https://github.com/rmcelreath/rethinking

library(ape)
library(geiger)
library(phytools)
library(dplyr)
library(tidyr)
library(rethinking)
library(cmdstanr)

# Keep this file (Code_IntersexDominance.Rmd) in a folder on your computer where you want all the output to be generated.
# This command will then tell R where to find the input files (if you choose the command versions operating locally) and where to save the output files.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Loading the data

# Version 1: from files locally on computer
# Keep this script file in the same folder as the three input files (population data, species variables, phylogeny)
fileorigin<-"local"
# Version 2: accessing the code and data directly from the online repository
fileorigin<-"github"


ifelse(fileorigin=="local",source("AnalysisCode/FemaleDominance_LoadData.r"),source("https://raw.githubusercontent.com/dieterlukas/primate_power/main/AnalysisCode/FemaleDominance_LoadData.r"))



# This should generate four files: the phylogenetic tree ('inputtree'), the population level data with the observations from the primary literature ('inputdata'), the species level data with predictors ('specieslevelpredictors'), and the data file combining these two datasets for the analyses ('combined'). It will also display in the console the list of variables.

```




### Descriptive summary of the data

## Occurrence of intersexual conflict

Intersexual conflicts are frequent, representing nearly half of all agonistic interactions among primates: mean percentage ±SD: 47.4±21.9

Intersexual conflicts are not more (or less) frequent when most dyads in the group are between the sexes: mean estimate -0.01, 89% HPDI -0.06 - +0.04


```{r p2 variation in intersexual conflicts, echo=FALSE}

mean(combined$perc_aggression_fm,na.rm=T)
sd(combined$perc_aggression_fm,na.rm=T)

dstan_conflicts<-select(combined,corrected_species_id,perc_aggression_fm,perc_fm_dyads)
dstan_conflicts<-dstan_conflicts[complete.cases(dstan_conflicts),]

  dat_list_conflicts <- list(
    perc_aggression_fm = as.integer(dstan_conflicts$perc_aggression_fm),
    perc_fm_dyads = standardize(dstan_conflicts$perc_fm_dyads),
    total = rep(100,nrow(dstan_conflicts))
    )
  
  m_conflicts <- ulam(
    alist(
      perc_aggression_fm ~ dbinom(total,p),
      logit(p) <-a +b*perc_fm_dyads,
      a ~dnorm(0,1),   
      b ~ dnorm(0,1)
    ) , data=dat_list_conflicts , chains=4 , cores=4 , log_lik=TRUE , cmdstan=T, messages=FALSE, refresh=0)

precis(m_conflicts)  
  
```



## Distribution of intersexual dominance

Running the following code will display tables and outputs directly in R with the descriptive statistics showing the distribution of intersexual dominance in the sample

```{r p1 variation in intersexual dominance, echo=FALSE}

### Distributinon of strict three way classification of intersexual dominance
# Total number of observations
nrow(combined)

# Total number of species
length(unique(combined$corrected_species_id))


# Distribution of intersexual dominance systems according to strict definition
# Occurrence of three types (1=strict male dominance, 2=no strict sex bias in dominance, 3=strict female dominance)
table(combined$strictfdom)

# Determine how many different species there are for each type of intersexual dominance
species_femdom_summarytable<-as.data.frame((combined %>% group_by(strictfdom) %>% summarize(unique(corrected_species_id)) ))

nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="1",])
nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="2",])
nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="3",])

# Determine the extent of variation within species

sum(table(combined$corrected_species_id)>1)

sum(table(combined$corrected_species_id)==1)


speciesvariation_femdom_summarytable<-as.data.frame((combined %>% group_by(corrected_species_id) %>% summarize(unique(strictfdom)) ))

# How many species have only a single classification
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==1)
# How many species have two classifications
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==2)
# How many species have three classifications
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==3)

combined[combined$corrected_species_id=="Miopithecus_talapoin",]


### Distribution of proportion of aggressive interactions won by females
# Total number of observations
nrow(combined[is.na(combined$perc_won_females)==FALSE,])

# Total number of species
length(unique(combined[is.na(combined$perc_won_females)==FALSE,]$corrected_species_id))

table(combined$perc_won_females)
hist(combined$perc_won_females)


# Checking for influence of captivity
dstan_captive <- combined[ complete.cases(combined$perc_won_females,combined$origin,combined$corrected_species_id),]

double<-as.data.frame(table(dstan_captive %>% group_by(corrected_species_id) %>% summarise(unique(origin))))
double<-double[double$Freq==1,]

table(double$corrected_species_id)[table(double$corrected_species_id)>1]

doublespecies<-names(table(double$corrected_species_id)[table(double$corrected_species_id)>1])

doubedstan_captive<-dstan_captive[dstan_captive$corrected_species_id %in% doublespecies,]

doubledstan_comp<-as.data.frame(doubedstan_captive %>% group_by(corrected_species_id,origin) %>% summarise(mean(perc_won_females)))

comparison<-NA
for (i in 1:7){
  
  comparison[i]<-doubledstan_comp[(2*i),]$`mean(perc_won_females)` - doubledstan_comp[(2*i-1),]$`mean(perc_won_females)`
}

comparison

```




## Phylogenetic signal of intersexual dominance

Running the following script will generate a csv output file with a table reporting the extent of the phylogenetic signal and their significance (results/FemaleDominance_PhylogeneticSignal.csv). In addition, it will generate two plots that show the estimated phylogenetic covariance based on the Bayesian model (figures/R_Figure_PhylogeneticCovariance...pdf).

```{r p1.4 phylogeny, echo=FALSE}

source("https://raw.githubusercontent.com/dieterlukas/primate_power/main/AnalysisCode/FemaleDominance_PhylogeneticSignal.r")

```






### Definition of functions to build statistical models to assess correlations with predictor variables

We link each of the predictor variables to the two outcome variables measuring intersexual dominance in Bayesian regression models. There are different models reflecting the different outcome variables (continuous versus categorical), whether the predictor variable is continuous or discrete, and whether the shared phylogenetic history among species is taken into account or not. For each of the unique combination of outcome variable and type of predictor variable, we built functions that can be called to automatically complete the respective set of analyses for a given predictor variable. These functions are loaded with the following script.


```{r functions, echo=FALSE}

source("https://raw.githubusercontent.com/dieterlukas/primate_power/main/AnalysisCode/FemaleDominance_FunctionsPairwise.r")


```






### Correlations with predictor variables

The following code runs the univariate analyses linking the predictor variables to the two outcome variables. The predictor variables are grouped according to the five hypotheses we are testing. The code will first generate a file with a table showing the sample sizes for the different predictor variables ("FemdomAnalyses_Samplesize.csv"). Next, it will run the correlations, with the estimates of the correlations of the two variables measuring intersexual dominance and the set of predictor variables provided in the output table "intersexualdominance_combinedresults.csv".


```{r p2.1 mating system, echo=FALSE}

# Running the analyses with the predictor variables for the female competition hypothesis

# a) social organization, b) fission-fusion, c) sexual size dimorphism, d) canine size dimorphism,  e) home range overlap, f) number of females
# bottom row: g) female evictions, h) population origin (ns), i) harshness (ns), j) rainfall seasonality (ns), k) rainfall unpredictability(ns), l) female infanticide (ns), m) relative canine size (ns), 

discrete_femalecompetition<-c("SocOrgPMK","fissionfusion","female_evictions","origin","female_infanticide")

continuous_femalecompetition<-c("SexualDimorphism_MaleWeight_over_FemaleWeight","CanineDimorphism","homerange_overlap","females","env_harshness","rainfall_annualvariation","rainfall_unpredictability","relative_femalecaninesize")


# Running the analyses with the predictor variables for the reproductive control hypothesis

# a) mating system, b) foraging location, c) sexual receptivity, d) adult sex ratio, e) male reproductive skew (ns), f) receptive synchrony (ns), g) relative testes mass (ns)

continuous_sexualselection<-c("sexualreceptivity_hours","sexratio","M_skew_index","Synchrony","relative_testes_mass")

discrete_sexualselection<-c("MatSysPMK","Strata_Wilman")


# Running the analyses with the predictor variables for the risk of offspring loss hypothesis

# a) parking offspring, b) short lactation periods, c) male infanticide, d) allomaternal care

continuous_offspringloss<-c("RelativeLactationDuration","AlloMaternalCare_continuous")

discrete_offspringloss<-c("Park","maleinfanticide","AlloMaternalCare")


# Running the analyses with the predictor variables for the female bonds hypothesis

# a) sex bias in dispersal, b) female relatedness (ns), c) female coalitions (ns)

continuous_femalesociality<-c("female_average_relatedness")

discrete_femalesociality<-c("sexbias_dispersal","jointaggression_females")


# Running the analyses with the predictor variables for the self organisation hypothesis

# a) adult sex ratio, b) male-male conflicts (ns), c) number of males (ns)

continuous_selforganisation<-c("sexratio","perc_aggression_mm","males")





# Combine all continuous predictors from the five hypotheses
allcontinuouspredictors<-c(continuous_femalecompetition,continuous_sexualselection,continuous_offspringloss,continuous_femalesociality,continuous_selforganisation)


# Combine all discrete predictors from the five hypotheses
alldiscretepredictors<-c(discrete_femalecompetition,discrete_sexualselection,discrete_offspringloss,discrete_femalesociality)





samplesizetable<-matrix(ncol=5,nrow=length(allcontinuouspredictors)+length(alldiscretepredictors))
samplesizetable<-as.data.frame(samplesizetable)
colnames(samplesizetable)<-c("predictorvariable","sample size categorical classification","number of species categorical classification", "sample size percentage fights won", "number of species percentage fights won")
count<-1

for(i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  colnames(combined)[colnames(combined) %in% continuouspredictor]<-c("continuouspredictor")
  samplesizetable[count,1]<-continuouspredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$continuouspredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$continuouspredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "continuouspredictor"]<-c(continuouspredictor)
  count<-count+1
}

for(i in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[i]
  colnames(combined)[colnames(combined) %in% discretepredictor]<-c("discretepredictor")
    samplesizetable[count,1]<-discretepredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$discretepredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$discretepredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "discretepredictor"]<-c(discretepredictor)
  count<-count+1
}

write.csv(samplesizetable,file="./results/FemdomAnalyses_Samplesize.csv")


for (k in 1:1){

# Start with the continuous predictors

for (i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  print(c("continuous",i,"of",length(allcontinuouspredictors),continuouspredictor))
  results<-run_analyses_continuouspredictor(continuouspredictor)
  ifelse(i==1,allresults<-results,allresults<-rbind(allresults,results))
  }


# Next with the discrete predictors

for (j in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[j]
  print(c("discrete",j,"of",length(alldiscretepredictors),discretepredictor))
  results<-run_analyses_discretepredictor(discretepredictor)
  ifelse(j==1,alldiscreteresults<-results,alldiscreteresults<-rbind(alldiscreteresults,results))
}

  colnames(alldiscreteresults)<-colnames(allresults)
  combinedresults<-rbind(allresults,alldiscreteresults)
  
  
combinedresults$association_present<-ifelse(combinedresults$'estimate lower'*combinedresults$'estimate upper'<0,"not","confident" )

# write.csv(combinedresults,file="./results/intersexualdominance_combinedresults.csv")

}


clip <- pipe("pbcopy", "w") 
write.table(allresults,sep=", ", file=clip)
close(clip)


```








We prepare the table for the results by grouping the variables according to the hypotheses.

```{r prepare pairwise tables, echo=FALSE}

source("https://raw.githubusercontent.com/dieterlukas/primate_power/main/AnalysisCode/FemaleDominance_PrepareTablesPairwise.r")

```





### Correlations with predictor variables while nesting within another variable

### Correlations with predictor variables

We repeat the analyses in subsets of the data. First, we split the sample into populations that are from lemur species versus non-lemur species, to assess whether a variable predicts variation in intersexual dominance beyond the clear split between lemurs and other species. Second, we split the sample into populations in which males are at least 10% larger than females, to assess whether additional factors shape intersexual dominance beyond the physical ability of males to outcompete females when they are much larger. The correlations of the two variables measuring intersexual dominance and the set of predictor variables are shown in the output table "intersexualdominance_combinedresults.csv".


```{r p2.1 mating system, echo=FALSE}

# Running the analyses with the predictor variables

original<-combined


discrete_femalecompetition<-c("SocOrgPMK","fissionfusion","female_evictions","origin","female_infanticide")
continuous_femalecompetition<-c("SexualDimorphism_MaleWeight_over_FemaleWeight","CanineDimorphism","homerange_overlap","females","env_harshness","rainfall_annualvariation","rainfall_unpredictability","relative_femalecaninesize")
continuous_sexualselection<-c("sexualreceptivity_hours","sexratio","M_skew_index","Synchrony","relative_testes_mass")
discrete_sexualselection<-c("MatSysPMK","Strata_Wilman")
continuous_offspringloss<-c("RelativeLactationDuration","AlloMaternalCare_continuous")
discrete_offspringloss<-c("Park","maleinfanticide","AlloMaternalCare")
continuous_femalesociality<-c("female_average_relatedness")
discrete_femalesociality<-c("sexbias_dispersal","jointaggression_females")
continuous_selforganisation<-c("sexratio","perc_aggression_mm","males")
allcontinuouspredictors<-unique(c(continuous_femalecompetition,continuous_sexualselection,continuous_femalesociality,continuous_selforganisation))
alldiscretepredictors<-c(discrete_femalecompetition,discrete_sexualselection,discrete_femalesociality)



for(nestingvariable in 1:2){

for(variant in 1:2){
  
  if(variant==1 & nestingvariable==1){
    combined<-original[original$lemur==0,]
    print("Lemurs")} 
  if(variant==2 & nestingvariable==1){combined<-original[original$lemur==1,]
  print("Not Lemurs")}
  
  if(variant==1 & nestingvariable==2){combined<-original[original$dimorphic==0,]
  print("Dimorphic")} 
  if(variant==2 & nestingvariable==2){combined<-original[original$dimorphic==1,]
  print("Not Dimorphic")}


samplesizetable<-matrix(ncol=5,nrow=length(allcontinuouspredictors)+length(alldiscretepredictors))
samplesizetable<-as.data.frame(samplesizetable)
colnames(samplesizetable)<-c("predictorvariable","sample size categorical classification","number of species categorical classification", "sample size percentage fights won", "number of species percentage fights won")
count<-1

for(i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  colnames(combined)[colnames(combined) %in% continuouspredictor]<-c("continuouspredictor")
  samplesizetable[count,1]<-continuouspredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$continuouspredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$continuouspredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "continuouspredictor"]<-c(continuouspredictor)
  count<-count+1
}

for(i in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[i]
  colnames(combined)[colnames(combined) %in% discretepredictor]<-c("discretepredictor")
    samplesizetable[count,1]<-discretepredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$discretepredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$discretepredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "discretepredictor"]<-c(discretepredictor)
  count<-count+1
}

if(variant==1 & nestingvariable==1){write.csv(samplesizetable,file="./results/subsets/FemdomAnalyses_NotLemur_Samplesizes.csv")}
if(variant==1 & nestingvariable==1){FemdomAnalyses_NotLemur_Samplesizes<-samplesizetable}
if(variant==2 & nestingvariable==1){write.csv(samplesizetable,file="./results/subsets/FemdomAnalyses_Lemur_Samplesizes.csv")}
if(variant==2 & nestingvariable==1){FemdomAnalyses_Lemur_Samplesizes<-samplesizetable}

if(variant==1 & nestingvariable==2){write.csv(samplesizetable,file="./results/subsets/FemdomAnalyses_NotDimorphic_Samplesizes.csv")}
if(variant==1 & nestingvariable==2){FemdomAnalyses_NotDimorphic_Samplesizes<-samplesizetable}
if(variant==2 & nestingvariable==2){write.csv(samplesizetable,file="./results/subsets/FemdomAnalyses_Dimorphic_Samplesizes.csv")}
if(variant==2 & nestingvariable==2){FemdomAnalyses_Dimorphic_Samplesizes<-samplesizetable}



for (k in 1:1){

# Start with the continuous predictors

for (i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  print(c("continuous",i,"of",length(allcontinuouspredictors),continuouspredictor))
  results<-run_analyses_continuouspredictor(continuouspredictor)
  ifelse(i==1,allresults<-results,allresults<-rbind(allresults,results))
  }


# Next with the discrete predictors

includeddiscretepredictors<-alldiscretepredictors  

for (k in 1:length(includeddiscretepredictors)){  
  discretepredictor<-alldiscretepredictors[k]
  colnames(combined)[colnames(combined) %in% discretepredictor]<-c("discretepredictor")
  dstan_strict <- combined[ complete.cases(combined$strictfdom,combined$discretepredictor,combined$corrected_species_id),]
  if(length(unique(dstan_strict$discretepredictor))==1){includeddiscretepredictors<-includeddiscretepredictors[-1]}
  colnames(combined)[colnames(combined) %in% "discretepredictor"]<-c(discretepredictor)
}
  
if(length(includeddiscretepredictors)>0){  
for (j in 1:length(includeddiscretepredictors)){
  discretepredictor<-includeddiscretepredictors[j]
  print(c("discrete",j,"of",length(includeddiscretepredictors),discretepredictor))
  results<-run_analyses_discretepredictor(discretepredictor)
  ifelse(j==1,alldiscreteresults<-results,alldiscreteresults<-rbind(alldiscreteresults,results))
}

  colnames(alldiscreteresults)<-colnames(allresults)
  combinedresults<-rbind(allresults,alldiscreteresults)
}
  
if(length(includeddiscretepredictors)==0){combinedresults<-allresults}
    
combinedresults$association_present<-ifelse(combinedresults$'estimate lower'*combinedresults$'estimate upper'<0,"not","confident" )

if(variant==1 & nestingvariable==1){write.csv(combinedresults,file="./results/subsets/intersexualdominance_notlemur_combinedresults.csv")}
if(variant==1 & nestingvariable==1){intersexualdominance_notlemur_combinedresults<-combinedresults}
if(variant==2 & nestingvariable==1){write.csv(combinedresults,file="./results/subsets/intersexualdominance_lemur_combinedresults.csv")}
if(variant==2 & nestingvariable==1){intersexualdominance_lemur_combinedresults<-combinedresults}

if(variant==1 & nestingvariable==2){write.csv(combinedresults,file="./results/subsets/intersexualdominance_notdimorphic_combinedresults.csv")}
if(variant==1 & nestingvariable==2){intersexualdominance_notdimorphic_combinedresults<-combinedresults}
if(variant==2 & nestingvariable==2){write.csv(combinedresults,file="./results/subsets/intersexualdominance_dimorphic_combinedresults.csv")}
if(variant==2 & nestingvariable==2){intersexualdominance_dimorphic_combinedresults<-combinedresults}

}

}

}

combined<-original




combinedresults<-combinedresults[combinedresults$phylogeny=="Yes",]
combinedresults<-combinedresults[combinedresults$outcome %in% c("perc_won", "strict three way"),]

combinedresults$estimate.lower<-round(combinedresults$'estimate lower',3)
combinedresults$estimate.upper<-round(combinedresults$'estimate upper',3)


femalecompetition<-c("SocOrgPMK","SexualDimorphism_MaleWeight_over_FemaleWeight","CanineDimorphism","homerange_overlap","fissionfusion","females","female_evictions","female_infanticide","relative_femalecaninesize","origin","env_harshness","rainfall_annualvariation","rainfall_unpredictability")
sexualselection<-c("MatSysPMK","Strata_Wilman","sexualreceptivity_hours","sexratio","M_skew_index","relative_testes_mass","Synchrony")
selforganisation<-c("sexratio","perc_aggression_mm","males")
socialbonds<-c("sexbias_dispersal","female_average_relatedness","jointaggression_females")
offspringloss<-c("RelativeLactationDuration","Park","AlloMaternalCare","AlloMaternalCare_continuous",             "maleinfanticide")

sorted_variables<-c(femalecompetition,sexualselection,offspringloss,selforganisation,socialbonds)

percwonresults<-combinedresults[combinedresults$outcome %in% "perc_won",]
threewayresults<-combinedresults[combinedresults$outcome %in% "strict three way",]

for(i in 1:length(sorted_variables)){
  currentrows<-agrep(sorted_variables[i],percwonresults$continuouspredictor,max.distance=0.1)
  if(i==1){sortedpercwonresults<-percwonresults[currentrows,]}else{sortedpercwonresults<-rbind(sortedpercwonresults,percwonresults[currentrows,])}
}

for(i in 1:length(sorted_variables)){
  currentrows<-agrep(sorted_variables[i],threewayresults$continuouspredictor,max.distance=0.1)
  if(i==1){sortedthreewayresults<-threewayresults[currentrows,]}else{sortedthreewayresults<-rbind(sortedthreewayresults,threewayresults[currentrows,])}
}

if(variant==1 & nestingvariable==1){
write.csv(sortedthreewayresults,file="./results/subsets/intersexualdominance_notlemur_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/subsets/intersexualdominance_notlemur_sortedpercwonresults.csv")
intersexualdominance_notlemur_sortedthreewayresults<-sortedthreewayresults
intersexualdominance_notlemur_sortedpercwonresults<-sortedpercwonresults
}

if(variant==2 & nestingvariable==1){
write.csv(sortedthreewayresults,file="./results/subsets/intersexualdominance_lemur_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/subsets/intersexualdominance_lemur_sortedpercwonresults.csv")
intersexualdominance_lemur_sortedthreewayresults<-sortedthreewayresults
intersexualdominance_lemur_sortedpercwonresults<-sortedpercwonresults  
}

if(variant==1 & nestingvariable==2){
write.csv(sortedthreewayresults,file="./results/subsets/intersexualdominance_notdimorphic_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/subsets/intersexualdominance_notdimorphic_sortedpercwonresults.csv")
intersexualdominance_notdimorphic_sortedthreewayresults<-sortedthreewayresults
intersexualdominance_notdimorphic_sortedpercwonresults<-sortedpercwonresults
}

if(variant==2 & nestingvariable==2){
write.csv(sortedthreewayresults,file="./results/subsets/intersexualdominance_dimorphic_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/subsets/intersexualdominance_dimorphic_sortedpercwonresults.csv")
intersexualdominance_dimorphic_sortedthreewayresults<-sortedthreewayresults
intersexualdominance_dimorphic_sortedpercwonresults<-sortedpercwonresults
}

```









Prepare the table for the results

```{r functions nested, echo=FALSE}

nestedresults<-read.csv("results/intersexualdominance_combinedresults_nestedbydimorphism_2023.csv")

nestedresults<-nestedresults[nestedresults$phylogeny=="yes",]

nestedresults$estimate.lower<-round(nestedresults$contrast.lower,3)
nestedresults$estimate.upper<-round(nestedresults$contrast.upper,3)

sexualselection_males<-c("MatSysPMK","SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism","sexratio","M_skew_index")
sexualselection_females<-c("Strata_Wilman","sexualreceptivity_hours","relative_testes_mass","ovulation_signs","Synchrony")
socialbonds<-c("SocOrgPMK","female_average_relatedness","jointaggression_females","sexbias_dispersal")
selforganisation<-c("sexratio","perc_aggression_mm","males")
femalecompetition_environment<-c("origin","env_harshness","rainfall_annualvariation","rainfall_unpredictability")
femalecompetition_social<-c("female_evictions","female_infanticide","r_seasonality_value","homerange_overlap","females","relative_femalecaninesize")

r_percwon<-nestedresults[nestedresults$outcome %in% "perc_won",]
r_strictdominance<-nestedresults[nestedresults$outcome %in% "strict",]

r_percwon_notdimorphic<-r_percwon[seq(from=1,to=69,by=2),]
r_percwon_notdimorphic<-r_percwon_notdimorphic[,c(3,7,8)]
r_percwon_dimorphic<-r_percwon[seq(from=2,to=70,by=2),]
r_percwon_dimorphic<-r_percwon_dimorphic[,c(3,7,8)]

r_strictdominance_notdimorphic<-r_strictdominance[seq(from=1,to=69,by=2),]
r_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[,c(3,7,8)]
r_strictdominance_dimorphic<-r_strictdominance[seq(from=2,to=70,by=2),]
r_strictdominance_dimorphic<-r_strictdominance_dimorphic[,c(3,7,8)]



r_sm_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_sf_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sm_percwon_notdimorphic,r_sf_percwon_notdimorphic),file="results/Table_S7.csv")

r_sb_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(socialbonds,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_so_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(selforganisation,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_fe_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_fs_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sb_percwon_notdimorphic,r_so_percwon_notdimorphic,r_fe_percwon_notdimorphic,r_fs_percwon_notdimorphic),file="results/Table_S8.csv")


r_sm_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_sf_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

write.csv(rbind(r_sm_percwon_dimorphic,r_sf_percwon_dimorphic),file="results/Table_S9.csv")


r_sb_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(socialbonds,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_so_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(selforganisation,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_fe_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_fs_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

write.csv(rbind(r_sb_percwon_dimorphic,r_so_percwon_dimorphic,r_fe_percwon_dimorphic,r_fs_percwon_dimorphic),file="results/Table_S10.csv")




r_sm_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_sf_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sm_strictdominance_notdimorphic,r_sf_strictdominance_notdimorphic),file="results/Table_S11.csv")


r_sb_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(socialbonds,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_so_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(selforganisation,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_fe_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_fs_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sb_strictdominance_notdimorphic,r_so_strictdominance_notdimorphic,r_fe_strictdominance_notdimorphic,r_fs_strictdominance_notdimorphic),file="results/Table_S12.csv")


r_sm_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_sf_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

write.csv(rbind(r_sm_strictdominance_dimorphic,r_sf_strictdominance_dimorphic),file="results/Table_S13.csv")

r_sb_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(socialbonds,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_so_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(selforganisation,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_fe_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_fs_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]


write.csv(rbind(r_sb_strictdominance_dimorphic,r_so_strictdominance_dimorphic,r_fe_strictdominance_dimorphic,r_fs_strictdominance_dimorphic),file="results/Table_S14.csv")


```




## Including Plots

The following code will generate the figures from the data and save them in the folder /figures.

```{r plots, echo=FALSE}

source("AnalysisCode/FemaleDominance_Figures.r")


```



### Analyses of variables explaining variation in intersexual dominance relations within species


For some species, we have data from multiple observations (different groups or different time points). The code in the following script compares the relationship between intersexual dominance and the three variables for which population-level information exists (number of females per group, number of males per group, adult sex ratio in group) either by averaging across the effect within each species or by averaging values within species to determine the effect across species. It needs to be run manually, it is not reported in the manuscript.

```{r variation, echo=FALSE}

source("AnalysisCode/FemaleDominance_VariationWithinSpecies.r")


```



### Path analyses to identify potential multi-variable interactions

In the discussion, we explore some of the potential interactions among the predictor variables. The code in the following script provides examples of the functions to include more than one predictor variables to assess their influence on the respective outcome variables measuring intersexual dominance.


```{r multifactor, echo=FALSE}

source("AnalysisCode/FemaleDominance_Multivariate.r")
```



