---
title: "IntersexDominance"
author: "Elise Huchard, Peter Kappeler, Claudia Fichtel, Nikos Smit, Dieter Lukas"
date: '`r Sys.Date()`'
output:
  html_document: 
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
    code_folding: hide 
  github_document: 
    toc: true
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
  md_document: 
    toc: true
---


### This file contains the R script with the analyses reported in the manuscript:
The evolution of male-female dominance relations in primate societies
Authors: Elise Huchard, Dieter Lukas, Nikolaos Smit, Claudia Fichtel, Peter M. Kappeler


## Abstract
In all but a few mammalian species, males were long believed to dominate females socially. However, recent studies indicate more variation, opening new possibilities for exploring the extent and drivers of sex biases in dominance relations. Here, we use comparative data from 226 populations of 117 primate species to show that male-female conflicts represent nearly half of all within-group conflicts, and that strict male dominance is found in <20% of study populations. Our analyses further indicate that female dominance emerges over evolutionary times as females face more reproductive competition and gain more reproductive control. These results shed new light on the evolutionary origins of power asymmetries between the sexes in non-human primates, with potential implications for understanding gender asymmetries in human societies.


# The script can be used to load the required data from the repository at http://github.com/dieterlukas/primate_power , run the analyses, and generate the figures.



```{r load data}



# Loading the relevant packages for tree manipulation (ape/geiger), data wrangling (dplyr), 
# and Bayesian analyses (rethinking/cmdstanr). To install these latter two packages, please see: https://github.com/rmcelreath/rethinking

library(ape)
library(geiger)
library(phytools)
library(dplyr)
library(rethinking)
library(cmdstanr)


# Loading the data

# Version 1: from files locally on computer
# Keep this script file in the same folder as the three input files (population data, species variables, phylogeny)
# This command will then tell R where to find these input files
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
source("AnalysisCode/FemaleDominance_LoadData.r")


# Version 2: accessing the code and data directly from the online repository
source("https://raw.githubusercontent.com/dieterlukas/primate_power/main/AnalysisCode/FemaleDominance_LoadData.r")

# This should generate four files: the phylogenetic tree ('inputtree'), the population level data with the observations from the primary literature ('inputdata'), the species level data with predictors ('specieslevelpredictors'), and the data file combining these two datasets for the analyes ('combined').

```




### Descriptive summary of the data

## Occurrence of intersexual conflict

Intersexual conflicts are frequent, representing nearly half of all agonistic interactions among primates: mean percentage ±SD: 48.1±21.5

Intersexual conflicts are not more (or less) frequent when most dyads in the group are between the sexes: mean estimate -0.01, 89% HPDI -0.06 - +0.04


```{r p2 variation in intersexual conflicts, echo=FALSE}

mean(combined$perc_aggression_fm,na.rm=T)
sd(combined$perc_aggression_fm,na.rm=T)

dstan_conflicts<-select(combined,corrected_species_id,perc_aggression_fm,perc_fm_dyads)
dstan_conflicts<-dstan_conflicts[complete.cases(dstan_conflicts),]

  dat_list_conflicts <- list(
    perc_aggression_fm = as.integer(dstan_conflicts$perc_aggression_fm),
    perc_fm_dyads = standardize(dstan_conflicts$perc_fm_dyads),
    total = rep(100,nrow(dstan_conflicts))
    )
  
  m_conflicts <- ulam(
    alist(
      perc_aggression_fm ~ dbinom(total,p),
      logit(p) <-a +b*perc_fm_dyads,
      a ~dnorm(0,1),   
      b ~ dnorm(0,1)
    ) , data=dat_list_conflicts , chains=4 , cores=4 , log_lik=TRUE , cmdstan=T, messages=FALSE, refresh=0)

precis(m_conflicts)  
  
```



## Distribution of intersexual dominance

Our final dataset consists of 226 observations from 117 primate species for the three-way categorical classification of intersexual dominance, and 128 observations from 77 primate species for the quantitative measure of the percentage of fights won by females. 

With the strict classification of intersexual dominance relations, 82 observations report that males dominate over females (47 different species), in 109 there is no clear sex-bias in the intersexual dominance interactions (65 different species), and in 35 females dominate over males (28 different species). 95 species in our sample are reported with only a single pattern of strict intersexual dominance, for 21 species we found reports of both no clear sex-bias in one more observations and either female or male dominance in other population(s), and in 1 species (Miopithecus talapoin) all three patterns of strict intrasexual dominance relations were described. 

Of the 128 quantitative measures of aggressive intersexual interactions, 23 show that females always win and another 21 show that males always win.


```{r p1 variation in intersexual dominance, echo=FALSE}

### Distributinon of strict three way classification of intersexual dominance
# Total number of observations
nrow(combined)

# Total number of species
length(unique(combined$corrected_species_id))


# Distribution of intersexual dominance systems according to strict definition
# Occurrence of three types (1=strict male dominance, 2=no strict sex bias in dominance, 3=strict female dominance)
table(combined$strictfdom)

# Determine how many different species there are for each type of intersexual dominance
species_femdom_summarytable<-as.data.frame((combined %>% group_by(strictfdom) %>% summarize(unique(corrected_species_id)) ))

nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="1",])
nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="2",])
nrow(species_femdom_summarytable[species_femdom_summarytable$strictfdom=="3",])

# Determine the extent of variation within species

speciesvariation_femdom_summarytable<-as.data.frame((combined %>% group_by(corrected_species_id) %>% summarize(unique(strictfdom)) ))

# How many species have only a single classification
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==1)
# How many species have two classifications
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==2)
# How many species have three classifications
sum(table(speciesvariation_femdom_summarytable$corrected_species_id)==3)

combined[combined$corrected_species_id=="Miopithecus_talapoin",]


### Distribution of proportion of aggressive interactions won by females
# Total number of observations
nrow(combined[is.na(combined$perc_won_females)==FALSE,])

# Total number of species
length(unique(combined[is.na(combined$perc_won_females)==FALSE,]$corrected_species_id))

table(combined$perc_won_females)
hist(combined$perc_won_females)

```




## Phylogenetic signal of intersexual dominance

All three variables of intersexual dominance show consistent, but limited phylogenetic signals: percentage fights won by females Blomberg's K 0.32 (p=0.001), strict intersexual dominance K 0.20 (p=0.001), relaxed intersexual dominance K 0.21 (p=0.001)

The phylogenetic covariance approach suggests that the percentage of fights that females can win changes in a somewhat Brownian fashion, with no very high covariances and a steady decline of covariance over phylogenetic distance.

In contrast, the three-way classification of whether females, males, or no sex are strictly dominant show extremely high phylogenetic signal at the level of Families, which drops to zero beyond that. For the three-way classification (female/co/male dominance), this reflects that states deeper in the tree cannot be accurately reconstructed (see also the figure).

These evolutionary patterns limit the ability to reconstruct co-evolutionary patterns for the two discrete traits. States in the the three-way, strict classification changes so frequently that there is high uncertainty around deeper parts of the tree.


```{r p1.4 phylogeny, echo=FALSE}

source("AnalysisCode/FemaleDominance_PhylogeneticSignal.r")

# Running this code will generate a csv output file with a table reporting the extent of the phylogenetic signal and their significance (results/FemaleDominance_PhylogeneticSignal.csv). In addition, it will generate two plots that show the estimated phylogenetic covariance based on the Bayesian model (figures/R_Figure_PhylogeneticCovariance...pdf).

```






### Definition of functions to build statistical models to assess correlations with predictor variables

We link each of the predictor variables to the three outcome variables measuring intersexual dominance in Bayesian regression models. There are different models reflecting the different outcome variables, whether the predictor variable is continuous or discrete, and whether the shared phylogenetic history among species is taken into account or not. For each of the unique combination of outcome variable and type of predictor variable, we built functions that can be called to automatically complete the respective set of analyses for a given predictor variable. 


```{r functions, echo=FALSE}

source("AnalysisCode/FemaleDominance_FunctionsPairwise.r")


```






### Correlations with predictor variables

The correlations of the two variables measuring intersexual dominance and the set of predictor variables are shown in the output table "intersexualdominance_combinedresults.csv".


```{r p2.1 mating system, echo=FALSE}

# Running the analyses with the predictor variables



allcontinuouspredictors<-c("sexratio", "SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism", "female_average_relatedness", "Synchrony", "r_seasonality_value", "M_skew_index", "env_harshness", "rainfall_unpredictability","rainfall_annualvariation","NDVI_Seasonality", "female_canine_height", "male_canine_height", "females","males","homerange_overlap","perc_aggression_mm","sexualreceptivity_hours","receptive_synchrony","body_mass","relative_testes_mass","relative_femalecaninesize")

continuous_sexualselection<-c("SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism","sexratio","M_skew_index","relative_testes_mass","sexualreceptivity_hours","Synchrony")
continuous_femalesociality<-c("female_average_relatedness")
continuous_selforganisation<-c("sexratio","perc_aggression_mm","males","females")
continuous_femalecompetition<-c("env_harshness","rainfall_unpredictability","rainfall_annualvariation","NDVI_Seasonality","r_seasonality_value","homerange_overlap","females","relative_femalecaninesize")

alldiscretepredictors<-c("SocOrgPMK","MatSysPMK","female_dispersal","male_dispersal","sexbias_dispersal","jointaggression_females","jointaggression_males","female_evictions","female_infanticide" ,"ovulation_signs","Strata_Wilman","origin","fissionfusion","between_groupconflict")

discrete_sexualselection<-c("MatSysPMK","Strata_Wilman","ovulation_signs")
discrete_selforganisation<-c("between_groupconflict")
discrete_femalesociality<-c("SocOrgPMK","jointaggression_females","sexbias_dispersal")
discrete_femalecompetition<-c("origin","female_evictions","female_infanticide","fissionfusion")

samplesizetable<-matrix(ncol=5,nrow=length(allcontinuouspredictors)+length(alldiscretepredictors))
samplesizetable<-as.data.frame(samplesizetable)
colnames(samplesizetable)<-c("predictorvariable","sample size categorical classification","number of species categorical classification", "sample size percentage fights won", "number of species percentage fights won")
count<-1

for(i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  colnames(combined)[colnames(combined) %in% continuouspredictor]<-c("continuouspredictor")
  samplesizetable[count,1]<-continuouspredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$continuouspredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$continuouspredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "continuouspredictor"]<-c(continuouspredictor)
  count<-count+1
}

for(i in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[i]
  colnames(combined)[colnames(combined) %in% discretepredictor]<-c("discretepredictor")
    samplesizetable[count,1]<-discretepredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$discretepredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$discretepredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "discretepredictor"]<-c(discretepredictor)
  count<-count+1
}

write.csv(samplesizetable,file="./results/FemdomAnalyses_Samplesizes.csv")


for (k in 1:1){

# Start with the continuous predictors

for (i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  print(c("continuous",i,"of",length(allcontinuouspredictors),continuouspredictor))
  results<-run_analyses_continuouspredictor(continuouspredictor)
  ifelse(i==1,allresults<-results,allresults<-rbind(allresults,results))
  }


# Next with the discrete predictors

for (j in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[j]
  print(c("discrete",j,"of",length(alldiscretepredictors),discretepredictor))
  results<-run_analyses_discretepredictor(discretepredictor)
  ifelse(j==1,alldiscreteresults<-results,alldiscreteresults<-rbind(alldiscreteresults,results))
}

  colnames(alldiscreteresults)<-colnames(allresults)
  combinedresults<-rbind(allresults,alldiscreteresults)
  
  
combinedresults$association_present<-ifelse(combinedresults$'estimate lower'*combinedresults$'estimate upper'<0,"not","confident" )

write.csv(combinedresults,file="./results/intersexualdominance_combinedresults.csv")

}



```




Prepare the table for the results

```{r functions nested, echo=FALSE}

combinedresults<-read.csv("results/intersexualdominance_combinedresults.csv")

combinedresults<-combinedresults[combinedresults$phylogeny=="Yes",]

combinedresults$estimate.lower<-round(combinedresults$estimate.lower,3)
combinedresults$estimate.upper<-round(combinedresults$estimate.upper,3)

sexualselection_males<-c("MatSysPMK","SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism","sexratio","M_skew_index")
sexualselection_females<-c("Strata_Wilman","sexualreceptivity_hours","relative_testes_mass","ovulation_signs","Synchrony")
socialbonds<-c("SocOrgPMK","female_average_relatedness","jointaggression_females","sexbias_dispersal")
selforganisation<-c("sexratio","perc_aggression_mm","males")
femalecompetition_environment<-c("origin","env_harshness","rainfall_annualvariation","rainfall_unpredictability")
femalecompetition_social<-c("female_evictions","female_infanticide","r_seasonality_value","homerange_overlap","females","relative_femalecaninesize","fissionfusion")

r_percwon<-combinedresults[combinedresults$outcome %in% "perc_won",]
r_strictdominance<-combinedresults[combinedresults$outcome %in% "strict three way",]
r_strictfemale<-combinedresults[combinedresults$outcome %in% "strict female dominance",]
r_strictmale<-combinedresults[combinedresults$outcome %in% "strict male dominance",]

r_sm_percwon<-r_percwon[unique(grep(paste(sexualselection_males,collapse="|"),r_percwon$continuouspredictor)),]
r_sm_percwon<-r_sm_percwon[c(5:10,2,3,1,4),]
r_sf_percwon<-r_percwon[unique(grep(paste(sexualselection_females,collapse="|"),r_percwon$continuouspredictor)),]
r_sf_percwon<-r_sf_percwon[c(7:9,2,3,4:6,1),]
r_sb_percwon<-r_percwon[unique(grep(paste(socialbonds,collapse="|"),r_percwon$continuouspredictor)),]
r_sb_percwon<-r_sb_percwon[c(2:4,1,8,5:7),]
r_so_percwon<-r_percwon[unique(grep(paste(selforganisation,collapse="|"),r_percwon$continuouspredictor)),]
r_so_percwon<-r_so_percwon[c(1,4,3),]
r_fe_percwon<-r_percwon[unique(grep(paste(femalecompetition_environment,collapse="|"),r_percwon$continuouspredictor)),]
r_fe_percwon<-r_fe_percwon[c(5,1,3,2),]
r_fs_percwon<-r_percwon[unique(grep(paste(femalecompetition_social,collapse="|"),r_percwon$continuouspredictor)),]
r_fs_percwon<-r_fs_percwon[c(6,7,1,8,3,2,4),]


r_sm_strictdominance<-r_strictdominance[unique(grep(paste(sexualselection_males,collapse="|"),r_strictdominance$continuouspredictor)),]
r_sm_strictdominance<-r_sm_strictdominance[c(5:10,2,3,1,4),]
r_sf_strictdominance<-r_strictdominance[unique(grep(paste(sexualselection_females,collapse="|"),r_strictdominance$continuouspredictor)),]
r_sf_strictdominance<-r_sf_strictdominance[c(7:9,2,3,4:6,1),]
r_sb_strictdominance<-r_strictdominance[unique(grep(paste(socialbonds,collapse="|"),r_strictdominance$continuouspredictor)),]
r_sb_strictdominance<-r_sb_strictdominance[c(2:4,1,8,5:7),]
r_so_strictdominance<-r_strictdominance[unique(grep(paste(selforganisation,collapse="|"),r_strictdominance$continuouspredictor)),]
r_so_strictdominance<-r_so_strictdominance[c(1,4,3),]
r_fe_strictdominance<-r_strictdominance[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictdominance$continuouspredictor)),]
r_fe_strictdominance<-r_fe_strictdominance[c(5,1,3,2),]
r_fs_strictdominance<-r_strictdominance[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictdominance$continuouspredictor)),]
r_fs_strictdominance<-r_fs_strictdominance[c(6,7,1,8,3,2,4),]



r_sm_strictfemale<-r_strictfemale[unique(grep(paste(sexualselection_males,collapse="|"),r_strictfemale$continuouspredictor)),]
r_sm_strictfemale<-r_sm_strictfemale[c(5:10,2,3,1,4),]
r_sf_strictfemale<-r_strictfemale[unique(grep(paste(sexualselection_females,collapse="|"),r_strictfemale$continuouspredictor)),]
r_sf_strictfemale<-r_sf_strictfemale[c(7:9,2,3,4:6,1),]
r_sb_strictfemale<-r_strictfemale[unique(grep(paste(socialbonds,collapse="|"),r_strictfemale$continuouspredictor)),]
r_sb_strictfemale<-r_sb_strictfemale[c(2:4,1,8,5:7),]
r_so_strictfemale<-r_strictfemale[unique(grep(paste(selforganisation,collapse="|"),r_strictfemale$continuouspredictor)),]
r_so_strictfemale<-r_so_strictfemale[c(1,4,3),]
r_fe_strictfemale<-r_strictfemale[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictfemale$continuouspredictor)),]
r_fe_strictfemale<-r_fe_strictfemale[c(5,1,3,2),]
r_fs_strictfemale<-r_strictfemale[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictfemale$continuouspredictor)),]
r_fs_strictfemale<-r_fs_strictfemale[c(6,7,1,8,3,2,4),]


r_sm_strictmale<-r_strictmale[unique(grep(paste(sexualselection_males,collapse="|"),r_strictmale$continuouspredictor)),]
r_sm_strictmale<-r_sm_strictmale[c(5:10,2,3,1,4),]
r_sf_strictmale<-r_strictmale[unique(grep(paste(sexualselection_females,collapse="|"),r_strictmale$continuouspredictor)),]
r_sf_strictmale<-r_sf_strictmale[c(7:9,2,3,4:6,1),]
r_sb_strictmale<-r_strictmale[unique(grep(paste(socialbonds,collapse="|"),r_strictmale$continuouspredictor)),]
r_sb_strictmale<-r_sb_strictmale[c(2:4,1,8,5:7),]
r_so_strictmale<-r_strictmale[unique(grep(paste(selforganisation,collapse="|"),r_strictmale$continuouspredictor)),]
r_so_strictmale<-r_so_strictmale[c(1,4,3),]
r_fe_strictmale<-r_strictmale[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictmale$continuouspredictor)),]
r_fe_strictmale<-r_fe_strictmale[c(5,1,3,2),]
r_fs_strictmale<-r_strictmale[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictmale$continuouspredictor)),]
r_fs_strictmale<-r_fs_strictmale[c(6,7,1,8,3,2,4),]


```





### Correlations with predictor variables while nesting within another variable

### Correlations with predictor variables

The correlations of the three variables measuring intersexual dominance and the set of predictor variables are shown in the output table "intersexualdominance_combinedresults.csv".


```{r p2.1 mating system, echo=FALSE}

# Running the analyses with the predictor variables

original<-combined

for(nestingvariable in 1:2){

for(variant in 1:2){
  
  if(variant==1 & nestingvariable==1){combined<-original[original$lemur==0,]} 
  if(variant==2 & nestingvariable==1){combined<-original[original$lemur==1,]}
  
  if(variant==1 & nestingvariable==2){combined<-original[original$dimorphic==0,]} 
  if(variant==2 & nestingvariable==2){combined<-original[original$dimorphic==1,]}

allcontinuouspredictors<-c("sexratio", "SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism", "female_average_relatedness", "Synchrony", "r_seasonality_value", "M_skew_index", "env_harshness", "rainfall_unpredictability","rainfall_annualvariation","NDVI_Seasonality", "female_canine_height", "male_canine_height", "females","males","homerange_overlap","perc_aggression_mm","sexualreceptivity_hours","receptive_synchrony","body_mass","relative_testes_mass","relative_femalecaninesize")


continuous_femalecompetition<-c("env_harshness","rainfall_unpredictability","rainfall_annualvariation","NDVI_Seasonality","r_seasonality_value","homerange_overlap","females","relative_femalecaninesize")
continuous_sexualselection<-c("SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism","sexratio","M_skew_index","relative_testes_mass","sexualreceptivity_hours","Synchrony")
continuous_femalesociality<-c("female_average_relatedness")
continuous_selforganisation<-c("sexratio","perc_aggression_mm","males","females")


alldiscretepredictors<-c("SocOrgPMK","MatSysPMK","female_dispersal","male_dispersal","sexbias_dispersal","jointaggression_females","jointaggression_males","female_evictions","female_infanticide" ,"ovulation_signs","Strata_Wilman","origin","fissionfusion","between_groupconflict")

discrete_femalecompetition<-c("origin","female_evictions","female_infanticide","fissionfusion")
discrete_sexualselection<-c("MatSysPMK","Strata_Wilman","ovulation_signs")
discrete_selforganisation<-c("between_groupconflict")
discrete_femalesociality<-c("SocOrgPMK","jointaggression_females","sexbias_dispersal")


samplesizetable<-matrix(ncol=5,nrow=length(allcontinuouspredictors)+length(alldiscretepredictors))
samplesizetable<-as.data.frame(samplesizetable)
colnames(samplesizetable)<-c("predictorvariable","sample size categorical classification","number of species categorical classification", "sample size percentage fights won", "number of species percentage fights won")
count<-1

for(i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  colnames(combined)[colnames(combined) %in% continuouspredictor]<-c("continuouspredictor")
  samplesizetable[count,1]<-continuouspredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$continuouspredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$continuouspredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$continuouspredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "continuouspredictor"]<-c(continuouspredictor)
  count<-count+1
}

for(i in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[i]
  colnames(combined)[colnames(combined) %in% discretepredictor]<-c("discretepredictor")
    samplesizetable[count,1]<-discretepredictor
  samplesizetable[count,2]<-nrow(combined[is.na(combined$discretepredictor)==F,])
  samplesizetable[count,3]<-length(unique(combined[is.na(combined$discretepredictor)==F,]$corrected_species_id))
  samplesizetable[count,4]<-nrow(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,])
  samplesizetable[count,5]<-length(unique(combined[is.na(combined$discretepredictor)==F & is.na(combined$perc_won_females)==F,]$corrected_species_id))
  colnames(combined)[colnames(combined) %in% "discretepredictor"]<-c(discretepredictor)
  count<-count+1
}

if(variant==1){write.csv(samplesizetable,file="./results/FemdomAnalyses_NotLemur_Samplesizes.csv")}
if(variant==2){write.csv(samplesizetable,file="./results/FemdomAnalyses_Lemur_Samplesizes.csv")}


for (k in 1:1){

# Start with the continuous predictors

for (i in 1:length(allcontinuouspredictors)){
  continuouspredictor<-allcontinuouspredictors[i]
  print(c("continuous",i,"of",length(allcontinuouspredictors),continuouspredictor))
  results<-run_analyses_continuouspredictor(continuouspredictor)
  ifelse(i==1,allresults<-results,allresults<-rbind(allresults,results))
  }


# Next with the discrete predictors

for (j in 1:length(alldiscretepredictors)){
  discretepredictor<-alldiscretepredictors[j]
  print(c("discrete",j,"of",length(alldiscretepredictors),discretepredictor))
  results<-run_analyses_discretepredictor(discretepredictor)
  ifelse(j==1,alldiscreteresults<-results,alldiscreteresults<-rbind(alldiscreteresults,results))
}

  colnames(alldiscreteresults)<-colnames(allresults)
  combinedresults<-rbind(allresults,alldiscreteresults)
  
  
combinedresults$association_present<-ifelse(combinedresults$'estimate lower'*combinedresults$'estimate upper'<0,"not","confident" )

if(variant==1){write.csv(combinedresults,file="./results/intersexualdominance_notlemur_combinedresults.csv")}
if(variant==2){write.csv(combinedresults,file="./results/intersexualdominance_lemur_combinedresults.csv")}

}


combinedresults<-combinedresults[combinedresults$phylogeny=="Yes",]
combinedresults<-combinedresults[combinedresults$outcome %in% c("perc_won", "strict three way"),]

combinedresults$estimate.lower<-round(combinedresults$'estimate lower',3)
combinedresults$estimate.upper<-round(combinedresults$'estimate upper',3)


femalecompetition<-c("SocOrgPMK","SexualDimorphism_MaleWeight_over_FemaleWeight","CanineDimorphism","homerange_overlap","fissionfusion","females","female_evictions","female_infanticide","relative_femalecaninesize","origin","env_harshness","rainfall_annualvariation","rainfall_unpredictability")
sexualselection<-c("MatSysPMK","Strata_Wilman","sexualreceptivity_hours","sexratio","M_skew_index","relative_testes_mass","Synchrony")
selforganisation<-c("sexratio","perc_aggression_mm","males")
socialbonds<-c("sexbias_dispersal","female_average_relatedness","jointaggression_females")

sorted_variables<-c(femalecompetition,sexualselection,selforganisation,socialbonds)

percwonresults<-combinedresults[combinedresults$outcome %in% "perc_won",]
threewayresults<-combinedresults[combinedresults$outcome %in% "strict three way",]

for(i in 1:length(sorted_variables)){
  currentrows<-agrep(sorted_variables[i],percwonresults$continuouspredictor,max.distance=0.1)
  if(i==1){sortedpercwonresults<-percwonresults[currentrows,]}else{sortedpercwonresults<-rbind(sortedpercwonresults,percwonresults[currentrows,])}
}

for(i in 1:length(sorted_variables)){
  currentrows<-agrep(sorted_variables[i],threewayresults$continuouspredictor,max.distance=0.1)
  if(i==1){sortedthreewayresults<-threewayresults[currentrows,]}else{sortedthreewayresults<-rbind(sortedthreewayresults,threewayresults[currentrows,])}
}

if(variant==1 & nestingvariable==1){
write.csv(sortedthreewayresults,file="intersexualdominance_notlemur_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="intersexualdominance_notlemur_sortedpercwonresults.csv")
}

if(variant==2 & nestingvariable==1){
write.csv(sortedthreewayresults,file="./results/intersexualdominance_lemur_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/intersexualdominance_lemur_sortedpercwonresults.csv")
}

if(variant==1 & nestingvariable==2){
write.csv(sortedthreewayresults,file="intersexualdominance_notdimorphic_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="intersexualdominance_notdimorphic_sortedpercwonresults.csv")
}

if(variant==2 & nestingvariable==2){
write.csv(sortedthreewayresults,file="./results/intersexualdominance_dimorphic_sortedthreewayresults.csv")
write.csv(sortedpercwonresults,file="./results/intersexualdominance_dimorphic_sortedpercwonresults.csv")
}

}
}
combined<-original

```









Prepare the table for the results

```{r functions nested, echo=FALSE}

nestedresults<-read.csv("results/intersexualdominance_combinedresults_nestedbydimorphism_2023.csv")

nestedresults<-nestedresults[nestedresults$phylogeny=="yes",]

nestedresults$estimate.lower<-round(nestedresults$contrast.lower,3)
nestedresults$estimate.upper<-round(nestedresults$contrast.upper,3)

sexualselection_males<-c("MatSysPMK","SexualDimorphism_MaleWeight_over_FemaleWeight", "CanineDimorphism","sexratio","M_skew_index")
sexualselection_females<-c("Strata_Wilman","sexualreceptivity_hours","relative_testes_mass","ovulation_signs","Synchrony")
socialbonds<-c("SocOrgPMK","female_average_relatedness","jointaggression_females","sexbias_dispersal")
selforganisation<-c("sexratio","perc_aggression_mm","males")
femalecompetition_environment<-c("origin","env_harshness","rainfall_annualvariation","rainfall_unpredictability")
femalecompetition_social<-c("female_evictions","female_infanticide","r_seasonality_value","homerange_overlap","females","relative_femalecaninesize")

r_percwon<-nestedresults[nestedresults$outcome %in% "perc_won",]
r_strictdominance<-nestedresults[nestedresults$outcome %in% "strict",]

r_percwon_notdimorphic<-r_percwon[seq(from=1,to=69,by=2),]
r_percwon_notdimorphic<-r_percwon_notdimorphic[,c(3,7,8)]
r_percwon_dimorphic<-r_percwon[seq(from=2,to=70,by=2),]
r_percwon_dimorphic<-r_percwon_dimorphic[,c(3,7,8)]

r_strictdominance_notdimorphic<-r_strictdominance[seq(from=1,to=69,by=2),]
r_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[,c(3,7,8)]
r_strictdominance_dimorphic<-r_strictdominance[seq(from=2,to=70,by=2),]
r_strictdominance_dimorphic<-r_strictdominance_dimorphic[,c(3,7,8)]



r_sm_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_sf_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sm_percwon_notdimorphic,r_sf_percwon_notdimorphic),file="results/Table_S7.csv")

r_sb_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(socialbonds,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_so_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(selforganisation,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_fe_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

r_fs_percwon_notdimorphic<-r_percwon_notdimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_percwon_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sb_percwon_notdimorphic,r_so_percwon_notdimorphic,r_fe_percwon_notdimorphic,r_fs_percwon_notdimorphic),file="results/Table_S8.csv")


r_sm_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_sf_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

write.csv(rbind(r_sm_percwon_dimorphic,r_sf_percwon_dimorphic),file="results/Table_S9.csv")


r_sb_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(socialbonds,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_so_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(selforganisation,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_fe_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

r_fs_percwon_dimorphic<-r_percwon_dimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_percwon_dimorphic$nestingvariable)),]

write.csv(rbind(r_sb_percwon_dimorphic,r_so_percwon_dimorphic,r_fe_percwon_dimorphic,r_fs_percwon_dimorphic),file="results/Table_S10.csv")




r_sm_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_sf_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sm_strictdominance_notdimorphic,r_sf_strictdominance_notdimorphic),file="results/Table_S11.csv")


r_sb_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(socialbonds,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_so_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(selforganisation,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_fe_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

r_fs_strictdominance_notdimorphic<-r_strictdominance_notdimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictdominance_notdimorphic$nestingvariable)),]

write.csv(rbind(r_sb_strictdominance_notdimorphic,r_so_strictdominance_notdimorphic,r_fe_strictdominance_notdimorphic,r_fs_strictdominance_notdimorphic),file="results/Table_S12.csv")


r_sm_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(sexualselection_males,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_sf_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(sexualselection_females,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

write.csv(rbind(r_sm_strictdominance_dimorphic,r_sf_strictdominance_dimorphic),file="results/Table_S13.csv")

r_sb_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(socialbonds,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_so_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(selforganisation,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_fe_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(femalecompetition_environment,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]

r_fs_strictdominance_dimorphic<-r_strictdominance_dimorphic[unique(grep(paste(femalecompetition_social,collapse="|"),r_strictdominance_dimorphic$nestingvariable)),]


write.csv(rbind(r_sb_strictdominance_dimorphic,r_so_strictdominance_dimorphic,r_fe_strictdominance_dimorphic,r_fs_strictdominance_dimorphic),file="results/Table_S14.csv")


```




## Including Plots

The following code will generate the figures from the data and save them in the folder /figures.

```{r plots, echo=FALSE}

source("AnalysisCode/FemaleDominance_Figures.r")


```



### Analyses of variables explaining variation in intersexual dominance relations within species


For some species, we have data from multiple observations (different groups or different time points). The code in the following script compares the relationship between intersexual dominance and the three variables for which population-level information exists (number of females per group, number of males per group, adult sex ratio in group) either by averaging across the effect within each species or by averaging values within species to determine the effect across species. It needs to be run manually, it is not reported in the manuscript.

```{r variation, echo=FALSE}

source("AnalysisCode/FemaleDominance_VariationWithinSpecies.r")


```



### Path analyses to identify potential multi-variable interactions

In the discussion, we explore some of the potential interactions among the predictor variables. The code in the following script provides examples of the functions to include more than one predictor variables to assess their influence on the respective outcome variables measuring intersexual dominance.


```{r multifactor, echo=FALSE}

source("AnalysisCode/FemaleDominance_Multivariate.r")
```



